---
title: "initial_results"
author: "AMANDA LIU"
date: "2017年11月15日"
output: html_document
---

Have some form of initial results. This could be a .Rmd document with a couple plots or .R script with a shiny app. The initial results should be short and to the point. Put the initial results into a folder called **initial_results**.

```{r}
require(readr)
require(dplyr)
require(stringr)
require(zipcode)
nypd.vechicle <- read_csv("NYPD_Motor_Vehicle_Collisions.csv", col_names = T)
summary(nypd.vechicle)
```

#Q1
What are the most dangerous/safe intersections in NYC?
```{r}
colnames(nypd.vechicle) <- str_replace_all(colnames(nypd.vechicle), " ", "_")
q1 <- nypd.vechicle %>% 
  mutate(num = 1,
         Num_Injureed = NUMBER_OF_PERSONS_INJURED + NUMBER_OF_PEDESTRIANS_INJURED + NUMBER_OF_CYCLIST_INJURED + NUMBER_OF_MOTORIST_INJURED,
         Num_Killed = NUMBER_OF_PERSONS_KILLED + NUMBER_OF_PEDESTRIANS_KILLED + NUMBER_OF_CYCLIST_KILLED + NUMBER_OF_MOTORIST_KILLED
         ) %>%
  filter(!is.na(CROSS_STREET_NAME)) %>%
  group_by(CROSS_STREET_NAME) %>%
  summarise(count = n(), 
            Num_Injureed = sum(Num_Injureed),
            Num_Killed = sum(Num_Killed))

knitr::kable(q1[q1$count == max(q1$count), ], 
             caption = "the most dangerous intersection in NYC")
knitr::kable(q1[q1$Num_Injureed == max(q1$Num_Injureed), ], 
             caption = "the most dangerous intersection in NYC")
knitr::kable(q1[q1$Num_Killed == max(q1$Num_Killed), ], 
             caption = "the most dangerous intersection in NYC")

safe <- q1[q1$count == min(q1$count) & q1$Num_Killed == 0 &
              q1$Num_Injureed == 0,]
knitr::kable(head(safe), caption = "the most safe intersection in NYC")
```

#Q2
Which type of vehicles (pickup truck, passenger, taxi, etc.) most frequently caused a collision?
```{r}
q2.1 <- nypd.vechicle %>%
  group_by(VEHICLE_TYPE_CODE_1) %>%
  summarise(n1 = n()) %>% 
  rename(VEHICLE_TYPE_CODE = VEHICLE_TYPE_CODE_1)
q2.2 <-  nypd.vechicle %>%
  group_by(VEHICLE_TYPE_CODE_2) %>%
  summarise(n2 = n()) %>%
  rename(VEHICLE_TYPE_CODE = VEHICLE_TYPE_CODE_2)
q2 <- inner_join(q2.1, q2.2) %>%
  mutate(n = n1 + n2) %>%
  filter(!VEHICLE_TYPE_CODE %in% c("UNKNOWN", "OTHER", NA)) %>%
  arrange(-n) %>%
  head(10)
knitr::kable(q2, caption = "The top 10 of vehicles most frequently caused a collision")
```

#Q3
What are the most popular contributing factors that led to a collision and is there a certain type of vehicle and/or area of NY associated with them?
```{r}
q3.1 <- nypd.vechicle %>%
  group_by(CONTRIBUTING_FACTOR_VEHICLE_1) %>%
  summarise(n1 = n()) %>% 
  rename(CONTRIBUTING_FACTOR_VEHICLE = CONTRIBUTING_FACTOR_VEHICLE_1)
q3.2 <-  nypd.vechicle %>%
  group_by(CONTRIBUTING_FACTOR_VEHICLE_2) %>%
  summarise(n2 = n()) %>%
  rename(CONTRIBUTING_FACTOR_VEHICLE = CONTRIBUTING_FACTOR_VEHICLE_2)
q3 <- inner_join(q3.1, q3.2) %>%
  mutate(n = n1 + n2) %>%
  filter(!CONTRIBUTING_FACTOR_VEHICLE %in% c("Unspecified", NA)) %>%
  arrange(-n) %>% 
  head(10)
knitr::kable(q3, caption = "The top 10 of popular contributing factors that led to a collision")
```

#Q4
What is the probability that people were injured or killed from the collision?
```{r}
q4 <- nypd.vechicle %>%
  mutate(Injureed_or_Killed = NUMBER_OF_PERSONS_INJURED + NUMBER_OF_PEDESTRIANS_INJURED + NUMBER_OF_CYCLIST_INJURED + NUMBER_OF_MOTORIST_INJURED +NUMBER_OF_PERSONS_KILLED + NUMBER_OF_PEDESTRIANS_KILLED + NUMBER_OF_CYCLIST_KILLED + NUMBER_OF_MOTORIST_KILLED
  ) %>%
  mutate(Injureed_or_Killed_Yes = ifelse(Injureed_or_Killed > 0, 1, 0))

prop <- mean(q4$Injureed_or_Killed_Yes)
knitr::kable(data.frame(prop),
             caption = "The probability that people were injured or killed from the collision")
  
```

#Q5
Are there any certain period that the collision often took place.
```{r}
# 30 minute
require(lubridate)
q5.1 <- nypd.vechicle %>%
  group_by(TIME) %>%
  summarise(n = n()) %>%
  mutate(time = hms(TIME),
         time = hour(time)*3600 + minute(time)*60 + second(time),
         times = ceiling(time/1800),
         time.end = time/1800
         )

q5.interval <- q5.1 %>%
  filter(times == time.end) %>%
  select(times, TIME)

q5.interval <- q5.interval %>%
  rename(begin.time = TIME) %>%
  inner_join(
    q5.interval %>%
      rename(end.time = TIME) %>%
  mutate(times = times - 1) 
  )


q5.2 <- q5.1 %>%
  group_by(times) %>%
  summarise(n = sum(n)) %>%
  inner_join(q5.interval) %>%
  arrange(-n) %>% 
  mutate(happn.times = n) %>%
  select(begin.time, end.time, happn.times) %>%
  head(10)
knitr::kable(q5.2, caption = "The top 10 of vehicles most frequently caused a collision")
```

