---
title: "Loops HW"
author: "AMANDA LIU"
date: "2017年10月16日"
output: html_document
---
#Amanda Liu
#PID:730042603

```{r}
library("tidyverse")
library("stringr")
```

Exercises
21.2.1 #1,2, 3ac 21.5.3 #1-4

21.2.1
#1
1.mean of every column in mtcars.
```{r}
means <- vector("double", ncol(mtcars))
names(means) <- names(mtcars)
for (i in names(mtcars)) {
  means[i] <- mean(mtcars[[i]])
}
means
```

2.Determine the type of each column in nycflights13::flights.
```{r}
data("flights", package = "nycflights13")
type <- vector("list", ncol(flights))
names(type) <- names(flights)
for (i in names(flights)) {
  type[[i]] <- class(flights[[i]])
}
type
```

3.Compute the number of unique values in each column of iris.
```{r}
data(iris)
uniq_value <- vector("double", ncol(iris))
names(uniq_value) <- names(iris)
for (i in names(iris)) {
  uniq_value[i] <- length(unique(iris[[i]]))
}
uniq_value
```

4.Generate 10 random normals for each of. Think about the output, sequence, and body before you start writing the loop.
```{r}
n <- 10
mu <- c(-10, 0, 10, 100)
normal <- vector("list", length(mu))
for (i in seq_along(normal)) {
  normal[[i]] <- rnorm(n, mean = mu[i])
}
normal
```


#2
Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:
```{r}
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}
out

#changed to
stringr::str_c(letters, collapse = "")

```

```{r}
x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))
sd

#changed to
sd(x)
```
```{r}
x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
out

#changed to
all.equal(cumsum(x),out)
```

#3
a.Write a for loop that prints() the lyrics to the children’s song “Alice the camel”.
c.Convert the song “99 bottles of beer on the wall” to a function. Generalise to any number of any vessel containing any liquid on any surface.

21.5.3 
#1
```{r}
#1
map_dbl(mtcars, mean)

#2
map(nycflights13::flights, class)

#3
map_int(iris, ~ length(unique(.)))

#4
map(c(-10, 0, 10, 100), rnorm, n = 10)
```

#2
How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?
```
We can use map_lgl and is.factor to determine whether or not it's a factor.
```
```{r}
map_lgl(mtcars, is.factor)
```

#3
What happens when you use the map functions on vectors that aren’t lists? What does map(1:5, runif) do? Why?
```{r}
map(1:5, runif)
```
```
"map" will be applied to every element in the vector.
```
#4
What does map(-2:2, rnorm, n = 5) do? Why? What does map_dbl(-2:2, rnorm, n = 5) do? Why?
```{r}
map(-2:2, rnorm, n = 5)
```
```
The sample size is 5 and it is from the normal distribution with means(-2,-1,0,1,2).
It returns each with a vector of length 5.
```
```
map_dbl(-2:2, rnorm, n = 5)
```
```
map_dbl will be an error. 
```
