---
title: "Data Analysis 1: dplyr and UNC departments"
author: "STOR 320"
output: html_document
---

# Amanda Liu PID: 730042603
Due **Monday 9/11/17** before the start of class. 

**You are to work alone on this assignment. (We will discuss.)**

Universities are typically opaque, bureaucratic institutions. Many public schools such as the University of North Carolina system make [data about university employees publically](http://www.newsobserver.com/news/databases/public-salaries/) available. Using this dataset what can we learn about the dynamics of the university system?


## dplyr

To download the RMarkdown file for this assignment, right-click on [.Rmd file for this assignment](https://www.dropbox.com/s/h18er3ap1v7w29k/Data%20Analysis%201.Rmd?dl=0), and select *open link in new tab*. Then proceed to download by saving the file to your *Working Directory*

The best way to to learn the dplyr package is to just use it a lot. Read through the following dplyr tutorial the complete questions in this .Rmd document. I suggest doing the coding in an R script then putting the answers into this documents.

## Instructions
- Please attempt these questions by Monday, September 11 at 4:40.  The graded version will be turned in on September 13, but it is essential that you make a serious attempt at Questions 1-10 by Monday.
- use `dplyr` functions to answer the following questions (base R will not be given credit) 
- submit a zipped folder containing the .Rmd file and the associated .html file to Sakai.

- When you are given directions, **return** something means just print it out. Similarly for questions **What is ...** just print out the answer.

Q: Return the sum of the first 10 positive integers?

Q: What is the sum of the first 10 positive integers?
```{r}
sum(1:10)
```

- **create** or **save** means save a variable

Create a variable `x` with the sum of the first 10 positive integers.
```{r}
x <- sum(1:10)
```



# The data

First download the data set from Ryan Thornburg's website. You may want to set `cache=TRUE` for this chunk so you don't have to download the data set every time you run the script.
```{r setup, message=FALSE, cache=FALSE}
library(tidyverse)
library(lubridate)

# Load the UNC departments data
data <- read_csv(url("http://ryanthornburg.com/wp-content/uploads/2015/05/UNC_Salares_NandO_2015-05-06.csv"))
```

Take a first look at the data
```{r}
colnames(data)

str(data)

head(data)
```

# Assignment


## Question 1
Return a data frame with columns: name, dept, age,totalsal 
```{r}
select(data, name, dept, age, totalsal)
```

Return a data frame with columns: statesal, nonstsal,totalsal using one of the `select_helpers` functions. If needed, google 'select_helpers in R'.
```{r}
select(data, contains("sal"))
```

## Question 2 
Rename the `fte` column to `fulltime`. Make sure this change is saved (i.e. `data <- ...`).
```{r}
data <- rename(data, fulltime = fte)
```

## Question 3 

What is the mean salary in the Neurosurgery department?

```{r}
group_by(data, dept) %>%summarise(mean(totalsal)) %>%filter(dept == "Neurosurgery")
```


Return a data frame with employee's in the Neurosurgery department making more than $500,000. Why might these professors be so well paid? 
```{r}
filter(data, dept == "Neurosurgery", totalsal >= 500000)

```
```
First, they are all full-time. And in general, people who works in neurosurgery earn a lot.
```
## Question 4 
What is the total amount that full time Dermatology employees get paid?


```{r}
filter(data, dept == "Dermatology", fulltime == 1)$totalsal %>% sum
```


## Question 5

Create a data frame called radio_dept whose rows are the employees from the Radiology department. 

- include only the following columns: name, position, age, nonstsal, totalsal. 
- order the employees by salary

First without pipes

```{r }
radio_dept <- arrange(select(filter(data, dept == "Radiology"), name, position, age, nonstsal, totalsal), desc(totalsal))
```


Next with pipes
```{r }
radio_dept <- filter(data, dept == "Radiology") %>% select(name, position, age, nonstsal, totalsal) %>% arrange(desc(totalsal))
```


Make a histogram of Radiology salaries
```{r }
ggplot(radio_dept) + geom_histogram(mapping = aes(x = totalsal),bins = 30) + xlab("Total salary")+ ylab("#Number") + ggtitle("Radiology department salaries")
```



## Question 6

Create a data frame called `dept_summary`  whose rows are the departments and  whose columns are: department size, mean department salary, median department salary, and maximum salary (using totalsal for salary). 

```{r }
dept_summary <- group_by(data, dept) %>%summarise(dept_size = n(), mean_dept_totalsal = mean(totalsal), med_dept_totalsal = median(totalsal), max_dept_totalsal = max(totalsal))
```

Order the departments by highest mean salary and print the 10 highest paid departments. 


```{r }
dept_summary %>%arrange(desc(mean_dept_totalsal)) %>%select(dept)
```

Order the departments by highest median salary and print the 10 highest paid departments. 

```{r }
dept_summary %>%arrange(desc(med_dept_totalsal)) %>%select(dept)
```

Why do these lists differ? If you were asked for the top 10 best paid departments at UNC which summary would you choose and why?

```
The lists differ because mean is different from median in general. Mean calculates the average of numbers and median refers to the midpoint of a group of data. In my opinion, median is more suitable to represent the best paid departments because the mean can more easily be influenced by the outliers.
```

## Question 7
How many departments have at least 10 employees?
```{r }
count(dept_summary, dept_size >= 10)
```


## Question 8
Which department hired the most people in 2010? Hint: you probably have to modify `hiredate`.
```{r }
mutate(data, hiredate = parse_datetime(hiredate, "%Y%m%d")) %>% filter(year(hiredate) == 2010) %>%group_by(dept) %>%summarise(N = n()) %>%arrange(desc(N))
```
```
In 2010, the department of medicine hired the most people.
```

## Question 9
Make a list of all the department names and sort this list alphabetically. What is the 42nd department in this list?
```{r }
select(data, dept) %>% distinct(dept) %>% arrange(dept) %>% slice()
select(data, dept) %>% distinct(dept) %>% arrange(dept) %>% slice(42)
```
```
The 42nd department in this list is "Soccer".
```


## Question 10
Plot number of people hired by the CS dept per year vs. year
```{r }
hires <- filter(data, dept %in% c("Computer Science", "Statistics and Operations Res", "Mathematics", "Biostatistics", "School of Info & Libr Science", "Physics-Astronomy")) %>%mutate(hiredate = parse_datetime(hiredate, "%Y%m%d"),yr = year(hiredate)) %>% group_by(yr, dept) %>%summarise(N = n()) %>%ungroup

filter(hires, dept == "Computer Science") %>%ggplot(aes(x = yr, y = N)) + theme_minimal() + geom_point(color = "#660066") + geom_line(color = "#FF66FF") + ylab("# of hired") + xlab("Year") + ggtitle("# of hired in the CS department")

```


Now add STOR, Math, Biostatistics, SILS and Physics to the above plot
```{r }
ggplot(hires, aes(x = yr, y = N, color = dept)) + theme_minimal() +
  geom_point()+ geom_line(alpha = 0.5 ) + scale_color_manual(name = "", values = c("#00FF00", "#FF66FF", "#3399CC", "#9933CC", "#FFFF00", "#333333")) + ylab("# of hired") + xlab("Year") + ggtitle("# of hired, different departments")
```


# Open question

Examine the relationship between age and salary at UNC. You can interpret this however you like e.g.

- age can mean number of years employed at the university
- compare and contrast different departments

Requirements

- hand in a separate .Rmd document 
    - load the data with read_csv(url(...)) like this script
- make 3 figures
    - a figure can be a visualization, print out of a data frame that you can created, the results of a statistical model/test (e.g. linear regression)
- display all code used
    - code should be commented
    - if you have a lot of functions you can put them in a separate .R script
- describe each figure
    - what is the figure (figure blah is a plot of foo vs. bar)
    - what is the upshot of the figure (figure blah shows that there is a strong relationship between foo and bar)
- write an explanation of your findings
    - a couple paragraphs
    - use proper grammar
    - your findings should be clear

# hints

Every question can be answered using: filter, arrange, select, mutate, summarize and group_by functions.

- here are two more dplyr tutorials: https://rpubs.com/justmarkham/dplyr-tutorial and https://www.coursera.org/learn/data-cleaning/home/week/3

- fte means full time employee