---
title: 'Part 2: predictive modeling'
output: html_document
---
#Amanda Liu PID:730042603

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```


__Part II__

First, load the data sets:

```{r}
library(tidyverse)

# data from 2011
hour11 <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/bikes_2011.csv')


# x data from 2012
hour12_x <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/bikes_12_x.csv')
```


Split data into train and test sets,  also, before modeling, we need to only use the variables which shown in the data 2012, so we need to drop variables such as 'casual' and transform some variables such as :


```{r}
set.seed(1)
use_variables = intersect(colnames(hour11), colnames( hour12_x))
hour11 = data.frame(hour11)
hour12_x = data.frame(hour12_x)
hour11 = hour11[,c(use_variables,"cnt")]
hour11 = hour11[, -c(2,4)]
hour12_x = hour12_x[,-c(2,4)]
  
for(i in c(2,3,4,6)) hour12_x[,i] = factor(hour12_x[,i])
for(i in c(2,3,4,6)) hour11[,i] = factor(hour11[,i])

train_index = sample(1:nrow(hour11), 0.8*nrow(hour11))
train = hour11[train_index,  ]
test = hour11[-train_index,  ]
```

Then we use two models, the first model is the full model with all of the predictors, for simple, i do not include high order terms such as interaction terms and quadratic terms, the second model is using BIC model selection to select from the full model, i train them on train set and test on test set and compare RMSE:


```{r}
full = lm(cnt ~ . - instant, data = train)
pred.full = predict(full, test)
pred.full = ifelse(pred.full < 0, 0, pred.full)
rmse1 = sqrt(mean((pred.full - test$cnt )^2))
rmse1

back = step(full, direction = "backward",k = log(nrow(train)), trace =0)


pred.back = predict(back, test)
pred.back = ifelse(pred.back < 0, 0,pred.back)

rmse2 = sqrt(mean((pred.back - test$cnt )^2))
rmse2

```

Note that, we replace negative values with 0s, as it can be found that using full model we obtain a lower RMSE, so we use full model to make predictions on 2012 data:


```{r}
Nfull = lm(cnt ~ . - instant, data = hour11)
# get predictions for 2012 data
pred_12 <- predict(full, newdata = hour12_x)
pred_12 = ifelse(pred_12  < 0, 0,pred_12 )
# add predictions to 2012 data frame
hour12_x <- hour12_x %>% 
            mutate(cnt_pred=pred_12)

# save the predictions
write_csv(hour12_x, 'bike_predictions_2012_ONYEN.csv')
```




